---
title: "Phenols"
author: "Lizzy Gibson"
date: "2/19/2019"
output:
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
options(scipen = 999)
library(haven)
library(tidyverse)
library(janitor)
library(reshape2)
library(broom)
library(tableone)
library(xtable)
```

## Read Data

```{r data_phenol}
library(MNdata)

phenol <- mn_edc %>% dplyr::select(sid, 11:18, 
                            grep("_detect", colnames(.))) %>% 
  dplyr::select(-10:-18) %>% 
  rename_all(~str_to_upper(str_remove(., "_sg_adj"))) %>% 
  drop_na()
```

### Clean Data

* Remove Non-Reported Values "NR"

* Assign > HS

Highest calibrator (ng/mL)
                                              
* BP-3 = 1000
* TCS = 1000
* 24-DCP = 100
* 25-DCP = 1000
* M-PB = 1000
* P-PB = 1000
* B-PB = 500

## % < LOD

LODs:

* benzophenone-3 (BP-3) = 0.4
* triclosan (TCS) = 2.3
* 2 4-dichlorophenol (24-DCP) = 0.2
* 2 5-dichlorophenol (25-DCP) = 0.2
* Methyl Paraben (M-PB) = 1.0
* Propyl Paraben (P-PB) = 0.2
* Butyl Paraben (B-PB) = 0.2

* BPA  = 0.4

BPA Value Labels:

* (0) "no flag"
* (1) "<LOD"
* (-9) "missing"

```{r}
lod_bpa <- phenol %>% dplyr::select(BPA_DETECT) %>% 
  table() %>% as.data.frame() %>% as_tibble() %>% 
  rename(lod = 1) %>% 
  spread(lod, Freq) %>% 
  mutate(Phenols = "BPA") %>% 
  rename(N_Detect = `0`, N_BLOD = '1') %>% 
  dplyr::select(Phenols, N_Detect, N_BLOD)
lod_bpa
```

* benzophenone-3 (BP-3) = 0.4
* triclosan (TCS) = 2.3
* 2 4-dichlorophenol (24-DCP) = 0.2
* 2 5-dichlorophenol (25-DCP) = 0.2
* Methyl Paraben (M-PB) = 1.0
* Propyl Paraben (P-PB) = 0.2
* Butyl Paraben (B-PB) = 0.2

```{r}
phenol_lod <- phenol %>% dplyr::select(SID, grep("DETECT", colnames(.)), -BPA_DETECT) %>%
  gather(Phenols, Detection, -SID) %>% 
  mutate(Detection = as.factor(Detection),
         Phenols = str_remove(Phenols, "_DETECT")) %>% 
  dplyr::select(-SID) %>% 
  group_by(Phenols) %>%
  count(Detection) %>% 
  mutate(Detection = case_when(Detection == "" ~ "N_Detect",
                          grepl("<LOD", Detection) ~ "N_BLOD",
                          grepl(">HS", Detection) ~ "N_Detect",
                          grepl("Result", Detection) ~ "Drop")) %>% 
  group_by(Phenols, Detection) %>% 
  summarise(N = sum(n)) %>% 
  spread(Detection, N) %>% 
  mutate(N_BLOD = as.integer(replace_na(N_BLOD, 0))) %>% 
  dplyr::select(Phenols, N_Detect, N_BLOD) %>% 
  bind_rows(lod_bpa)
  
detection <- phenol_lod %>% 
  mutate(P_Detect = round(N_Detect/(N_BLOD + N_Detect), 2),
         P_BLOD = round(N_BLOD/(N_BLOD + N_Detect), 2))
detection
```

## Concentrations

```{r}
conc = phenol %>% dplyr::select(2:9) %>% 
  pivot_longer(DCP_24:BPA,
               names_to = "Phenols") %>% 
  group_by(Phenols) %>% 
  summarise(qs = quantile(value, c(0.25, 0.5, 0.75), na.rm = TRUE), prob = c("Q25", "Median", "Q75"),
            Mean = mean(value),
            Stdev = sd(value),
            Max = max(value)) %>% 
  pivot_wider(names_from = "prob",
              values_from = "qs") %>% 
  dplyr::select(Phenols, Mean, Stdev, Q25, Median, Q75, Max)
conc
```

## Table

```{r}
conc_table <- left_join(detection, conc, by = "Phenols") %>%
  ungroup() %>% 
  dplyr::select(Phenols, `<LOD` = N_BLOD, Detected = N_Detect, 
                Prop_Detected = P_Detect, everything(), -P_BLOD) %>% 
  mutate_if(is.numeric, round, digits = 2)

conc_table %>% knitr::kable()

#xtable(conc_table)
```

## Heat Map

```{r}
heat <- phenol %>% dplyr::select(2:9)

cormat <- round(cor(heat, use = "pairwise.complete.obs", method = c("spearman")),2)

melted_cormat <- melt(cormat) %>% rename(Correlation = value)

ggplot(data = melted_cormat, aes(x = Var1, y = Var2)) + 
  geom_tile(aes(fill = Correlation), colour = "white") + 
  scale_fill_gradient2(low = "#00BFC4", mid = "white", high = "#F8766D", 
                       midpoint = 0, 
                      na.value = "transparent", limits = c(-1, 1)) +
  labs(x = "", y = "", title = "") + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom")
```

