---
title: "Phthalates"
author: "Lizzy Gibson"
date: "2/19/2019"
output:
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
options(scipen = 999)
library(haven)
library(tidyverse)
library(janitor)
library(reshape2)
library(broom)
library(tableone)
library(xtable)
```

## Read Data

```{r pcb}
library(MNdata)

pht <- mn_edc %>% dplyr::select(1:10, grep("detect", colnames(.))) %>%
  dplyr::select(-grep("_m2", colnames(.))) %>%
  dplyr::select(1:19) %>% 
  rename_all(~str_remove(., "_sg_adj")) %>% 
  drop_na()
pht
```

## % < LOD

LODs:

* MECPP = 0.6
* MEHHP = 0.7
* MEOHP = 0.7
* MCPP = 0.2
* MiBP = 0.3
* MnBP = 0.6
* MBzP = 0.216
* MEP = 0.528
* MEHP = 1.2
* MCOP = 0.7
* MCNP = 0.6

Value Labels:

(1) "below LOD"
(0) "no flag"
(-9) "missing"
(2) "mean of ufiles 2&3, no flag"
(3) "mean of ufiles 3 & 8, no flag"
(4) "mean of ufiles revised & 3, no flag"

Remove (-9) missing because of QC issues or carryover.

```{r}
lod <- pht %>% 
  dplyr::select(grep("detect", colnames(.))) %>%
  gather(pht, level) %>% 
  group_by(pht, level) %>% 
  summarize(n = n()) %>% 
  spread(level, n) %>% 
  mutate_all(~replace_na(., 0))
lod
# Number of obs with each flag for each pht

detected <- lod %>% 
  mutate(Prop_Detected = `0`/(`0` + `1`)) %>%
  ungroup() %>% 
  dplyr::select(pht, `<LOD` = `1`, Detected = `0`, Prop_Detected) %>% 
  mutate(pht = str_remove(pht, "_detect"),
         pht = str_to_upper(pht)) %>% 
  rename(PHT = pht)
detected
```

## Concentrations

** **Need to handle <LOD** **
```{r}
conc = pht %>% dplyr::select(2:10) %>% 
  pivot_longer(MECPP:MEHP,
               names_to = "PHT") %>% 
  group_by(PHT) %>% 
  summarise(qs = quantile(value, c(0.25, 0.5, 0.75), na.rm = TRUE), prob = c("Q25", "Median", "Q75"),
            Mean = mean(value),
            Stdev = sd(value),
            Max = max(value)) %>% 
  pivot_wider(names_from = "prob",
              values_from = "qs") %>% 
  dplyr::select(PHT, Mean, Stdev, Q25, Median, Q75, Max)
```

## Table

```{r}
pht_conc <- left_join(detected, conc, by = c("PHT")) %>%
  mutate_if(is.numeric, round, digits = 2)
pht_conc

pht_conc  %>% knitr::kable()

#xtable(pht_conc)
```

