---
title: "Phthalate & phenol exposure patterns"
subtitle: "& child intelligence in M&N cohort"
author: "Lizzy Gibson"
date: "3/16/2021"
output:
  html_document:
    toc: yes
    toc_float: yes

---

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align='center', cache=TRUE,
                      autodep = TRUE)

options(mc.cores = parallel::detectCores())
source("./source/packages.R")
# source("./source/read_data.R")
```

## Mothers & Newborns

* 5 phenols  
* 3 parabens  
* 9 phthalates  

```{r, dat}
# Robbie: Perhaps one time we can go through this section on Zoom in a live session? I am finding it difficult to interpret abstractly... particularly what the results of model.matrix look like
# Lizzy: sure thing!

# Robbie: Is the following (alpha_w, beta_w, alpha_a), beta_a) derived from a step with BN2MF?
# Lizzy: Yes, BN2MF outputs Gamma dist alpha and beta for all three solution matrices, W, a, and H

# UNSTANDARDIZED GAMMA*GAMMA
alpha_w <- readMat(here::here("./Data/mn2_Walpha.mat"))[[1]] %>% as_tibble() %>% rename(walpha1 = V1, walpha2 = V2)
beta_w <- readMat(here::here("./Data/mn2_Wbeta.mat"))[[1]] %>% as_tibble() %>% rename(wbeta1 = V1, wbeta2 = V2)

alpha_a <- readMat(here::here("./Data/mn2_Aalpha.mat"))[[1]] %>% as_tibble() %>% rename(aalpha1 = V1, aalpha2 = V2)
beta_a <- readMat(here::here("./Data/mn2_Abeta.mat"))[[1]] %>% as_tibble() %>% rename(abeta1 = V1, abeta2 = V2)

bayes = bind_cols(ppp_cov, ewa, alpha_w, beta_w) %>%
        dplyr::select(-GEST, -ETH) %>%
        rename(HOME_SCORE = HOMETOT) %>%
        filter(P1 != max(P1)) %>%
        filter(P1 != max(P1)) %>%  # removes two outliers
        drop_na() %>%
        mutate_at(vars(c(HOME_SCORE, M_AGE, M_IQ)), scale) %>%
        mutate_if(is.character, as_factor)

# # For stan model ####
N = nrow(bayes)
C = 7 # number of covariates
x = model.matrix(WISC ~ M_AGE + M_EDU + MARITAL_STATUS + HOME_SCORE + M_IQ + ALCOHOL + mat_hard,
                 data = bayes)[,-1]
head(x)
# Lizzy: model.matrix basically makes dummies (if applicable) and creates a matrix of the included covariates
# [,-1] at the end gets rid of the intercept value that model.matrix creates
sex = model.matrix(WISC ~ SEX, data = bayes)[,-1]
y = bayes$WISC

# # Predictors of interest

# ### Unscaled Gamma
Walpha1 = bayes$walpha1

Aalpha1 = alpha_a[[1]]
Abeta1  = beta_a[[1]]

Walpha2 = bayes$walpha2

Aalpha2 = alpha_a[[2]]
Abeta2  = beta_a[[2]]

### Scaled and Standardized Gamma dist for W
# for a Gamma distribution, cX ~ Gamma(alpha, beta/c)
# don't scale alpha, only scale beta
# 1. Scale W Gamma distribution by multiplying the distribution by the normalization constant used to normalize H loadings (ie make them sum to 1)
## this is the 17.#### number
## It came from the MATLAB output
# 2. Next scale W to have standard deviation = 1
## standard deviation of EWA = 4.4074 for pattern 1 and 3.3636 for pattern 2 (after first scaling)
## it also came from MATLAB output
## X/c ~ Gamma(alpha, beta*c)

Wbeta1  = (bayes$wbeta1/17.3043)*4.4074
Wbeta2  = (bayes$wbeta2/17.2979)*3.3636

# don't scale a only scale W
```

```{r, standat}
# Robbie: So Stan always takes in a list for input data?
# Lizzy: Yes! At least, the R-Stan interface does it that way.
p1_dat = list(N = N,
                  C = C,
                  x = x,
                  y = y,
                  sex = sex,
                  alpha_w = Walpha1,
                  beta_w  = Wbeta1,
                  alpha_a = Aalpha1,
                  beta_a  = Abeta1)

p2_dat = list(N = N,
                  C = C,
                  x = x,
                  y = y,
                  sex = sex,
                  alpha_w = Walpha2,
                  beta_w  = Wbeta2,
                  alpha_a = Aalpha2,
                  beta_a  = Abeta2)
```

## Bayesian models

* this is only non-informative models
# Robbie: It looks like all the models are here. Is the second point above needed?
# Lizzy: Deleted!

```{r}
p1_non <- stan(
  file = "./Stan/models/bn2mf_score_reg_int_gamma.stan",  # Stan program
    data = p1_dat,               # named list of data
    chains = 4,                   # number of Markov chains
    warmup = 1000,                # number of warmup iterations per chain
    iter = 5000,                  # total number of iterations per chain
    cores = 2)                    # number of cores
# save(p1_non, file = "./Stan/fits/pattern1_non.rda")
 
p2_non <- stan(
  file = "./Stan/models/bn2mf_score_reg_int_gamma.stan",  # Stan program
    data = p2_dat,               # named list of data
    chains = 4,                   # number of Markov chains
    warmup = 1000,                # number of warmup iterations per chain
    iter = 5000,                  # total number of iterations per chain
    cores = 2)                    # number of cores
 # save(p2_non, file = "./Stan/fits/pattern2_non.rda")

# p1_inf <- stan(
#   file = "./Stan/models/bn2mf_score_reg_int_infp1.stan",  # Stan program
#     data = p1_dat,               # named list of data
#     chains = 4,                   # number of Markov chains
#     warmup = 1000,                # number of warmup iterations per chain
#     iter = 5000,                  # total number of iterations per chain
#     cores = 2)                    # number of cores
# save(p1_inf, file = "./Stan/fits/pattern1_inform.rda")

# p2_inf <- stan(
#   file = "./Stan/models/bn2mf_score_reg_int_infp2.stan",  # Stan program
#     data = p2_dat,               # named list of data
#     chains = 4,                   # number of Markov chains
#     warmup = 1000,                # number of warmup iterations per chain
#     iter = 5000,                  # total number of iterations per chain
#     cores = 2)                    # number of cores
# save(p2_inf, file = "./Stan/fits/pattern2_inform.rda")

# p1_hybrid <- stan(
#   file = "./Stan/models/bn2mf_score_reg_int_infp1_nonc.stan",  # Stan program
#     data = p1_dat,               # named list of data
#     chains = 4,                   # number of Markov chains
#     warmup = 1000,                # number of warmup iterations per chain
#     iter = 5000,                  # total number of iterations per chain
#     cores = 2)                    # number of cores
# save(p1_hybrid, file = "./Stan/fits/pattern1_hybrid.rda")

# p2_hybrid <- stan(
#   file = "./Stan/models/bn2mf_score_reg_int_infp2_nonc.stan",  # Stan program
#     data = p2_dat,               # named list of data
#     chains = 4,                   # number of Markov chains
#     warmup = 1000,                # number of warmup iterations per chain
#     iter = 5000,                  # total number of iterations per chain
#     cores = 2)                    # number of cores
# save(p2_hybrid, file = "./Stan/fits/pattern2_hybrid.rda")
```

**You have these .rda files, so you can run from here down.**
```{r, stan}
load("./Stan/fits/pattern1_non.rda")
load("./Stan/fits/pattern2_non.rda")

load("./Stan/fits/pattern1_inform.rda")
load("./Stan/fits/pattern2_inform.rda")

load("./Stan/fits/pattern1_hybrid.rda")
load("./Stan/fits/pattern2_hybrid.rda")
```

```{r}
trace = c("beta_int", "beta_sex", "beta_p")

stan_trace(p1_non, pars = trace, nrow = 3)
```

### Summary

#### Pattern 1

##### Non-informative
```{r, p1non}
params = names(p1_non)[grep("(alpha|beta)", names(p1_non))]
# print(p1_non, params)
add_ci4int_bayes(p1_non)
```

##### Informative Pattern
```{r, p1inf}
# print(p1_inf, params)
add_ci4int_bayes(p1_inf)
```

##### Informative Everything
```{r, p1inf3}
# print(p1_hybrid, params)
add_ci4int_bayes(p1_hybrid)
```

##### All

```{r, all1}
add_ci4int_bayes(p1_non)[,1:2] %>% 
  rename(non_inf = mean) %>% 
  bind_cols(., inf_pat =add_ci4int_bayes(p1_inf)[,2]) %>% 
  bind_cols(., inf_cov = add_ci4int_bayes(p1_hybrid)[,2])
```

#### Pattern 2

##### Non-informative
```{r, p2non}
params = names(p2_non)[grep("(alpha|beta)", names(p2_non))]
# print(p2_non, params)
add_ci4int_bayes(p2_non)
```

##### Informative Pattern
```{r, p2inf}
# print(p2_inf, params)
add_ci4int_bayes(p2_inf)
```

##### Informative Everything
```{r, p1inf1}
# print(p2_hybrid, params)
add_ci4int_bayes(p2_hybrid)
```

#### All

```{r, all2}
add_ci4int_bayes(p2_non)[,1:2] %>% 
  rename(non_inf = mean) %>% 
  bind_cols(., inf_pat =add_ci4int_bayes(p2_inf)[,2]) %>% 
  bind_cols(., inf_ncov = add_ci4int_bayes(p2_hybrid)[,2])
```

### Viz

#### Pattern 1

##### Non-informative
```{r}
ext_p1 <- extract(p1_non)
p1non_coef = cbind(ext_p1$beta_int, ext_p1$beta_sex, ext_p1$beta_p) %>% 
  as_tibble() %>% 
  rename(int=1, sex=2, p_male=3) %>% 
             mutate(p_female = p_male + int) %>% 
  dplyr::select(-sex)

mcmc_areas(p1non_coef,
           prob = 0.95) + ggtitle("Posterior distributions",
                                  "with medians and 95% intervals") +
  scale_y_discrete(labels=c(
  "int" = "Interaction", 
  "p_male" = "P1 in males",
  "p_female" = "P1 in females")) + xlim(c(-15,15))
```

##### Informative Pattern
```{r}
ext_p1i = extract(p1_hybrid)
p1i_coef = cbind(ext_p1i$beta_int, ext_p1i$beta_sex, ext_p1i$beta_p) %>% 
  as_tibble() %>% 
  rename(int=1, sex=2, p_male=3) %>% 
             mutate(p_female = p_male + int) %>% 
  dplyr::select(-sex)

mcmc_areas(p1i_coef,
           prob = 0.95) + ggtitle("Posterior distributions",
                                  "with medians and 95% intervals") +
  scale_y_discrete(labels=c(
  "int" = "Interaction", 
  "p_male" = "P1 in males",
  "p_female" = "P1 in females")) + xlim(c(-15,15))
```

##### Informative Everything
```{r}
ext_p1inc = extract(p1_inf)
p1i_coef_nc = cbind(ext_p1inc$beta_int, ext_p1inc$beta_sex, ext_p1inc$beta_p) %>% 
  as_tibble() %>% 
  rename(int=1, sex=2, p_male=3) %>% 
             mutate(p_female = p_male + int) %>% 
  dplyr::select(-sex)

mcmc_areas(p1i_coef_nc,
           prob = 0.95) + ggtitle("Posterior distributions",
                                  "with medians and 95% intervals") +
  scale_y_discrete(labels=c(
  "int" = "Interaction", 
  "p_male" = "P1 in males",
  "p_female" = "P1 in females")) + xlim(c(-15,15))
```

#### Pattern 2

##### Non-informative
```{r}
ext_p2 = extract(p2_non)
p2non_coef = cbind(ext_p2$beta_int, ext_p2$beta_sex, ext_p2$beta_p) %>% 
  as_tibble() %>% 
  rename(int=1, sex=2, p_male=3) %>% 
             mutate(p_female = p_male + int) %>% 
  dplyr::select(-sex)

mcmc_areas(p2non_coef,
           prob = 0.95) + ggtitle("Posterior distributions",
                                  "with medians and 95% intervals") +
  scale_y_discrete(labels=c(
  "int" = "Interaction", 
  "p_male" = "P2 in males",
  "p_female" = "P2 in females")) + xlim(c(-12,12))
```

##### Informative Pattern
```{r}
ext_p2i = extract(p2_hybrid)
p2i_coef = cbind(ext_p2i$beta_int, ext_p2i$beta_sex, ext_p2i$beta_p) %>% 
  as_tibble() %>% 
  rename(int=1, sex=2, p_male=3) %>% 
             mutate(p_female = p_male + int) %>% 
  dplyr::select(-sex)

mcmc_areas(p2i_coef,
           prob = 0.95) + ggtitle("Posterior distributions",
                                  "with medians and 95% intervals") +
  scale_y_discrete(labels=c(
  "int" = "Interaction", 
  "p_male" = "P2 in males",
  "p_female" = "P2 in females")) + xlim(c(-12,12))
```

##### Informative Everything
```{r}
ext_p2inc = extract(p2_inf)
p2i_coef_nc = cbind(ext_p2inc$beta_int, ext_p2inc$beta_sex, ext_p2inc$beta_p) %>% 
  as_tibble() %>% 
  rename(int=1, sex=2, p_male=3) %>% 
             mutate(p_female = p_male + int) %>% 
  dplyr::select(-sex)

mcmc_areas(p2i_coef_nc,
           prob = 0.95) + ggtitle("Posterior distributions",
                                  "with medians and 95% intervals") +
  scale_y_discrete(labels=c(
  "int" = "Interaction", 
  "p_male" = "P2 in males",
  "p_female" = "P2 in females")) + xlim(c(-12,12))
```

## Table 3
```{r}
bayesp1 = add_ci4int_bayes(p1_non)[c(9,11,12),]%>% mutate(pattern = "Pattern 1")
bayesp2 = add_ci4int_bayes(p2_non)[c(9,11:12),]%>% mutate(pattern = "Pattern 2")

bayes_table = bind_rows(bayesp1, bayesp2)[,c(1,7, 5, 9:10)] %>% as_tibble() %>% 
            rename(term = beta, estimate = `50%`) %>% 
            rename_all(~str_replace(., "%", " %")) %>% 
            mutate(model = "Bayesian") %>% 
  dplyr::select(model, pattern, everything()) %>% 
  arrange(model, pattern, term) %>% 
  mutate_if(is.numeric, round, 1) %>% 
  mutate(`95% Confidence Interval` = str_c("(", `2.5 %`, ", ", `97.5 %`, ")")) %>% 
  mutate(term = case_when(term == "sex*p" | grepl("SEX", term) ~ " interaction term",
                          TRUE ~ term),
         term = str_remove(term, "(P1|P2|p)"),
         term = str_c(pattern, term)) %>% 
    dplyr::select(-`2.5 %`, -`97.5 %`, -pattern)
bayes_table

# health model table
# stargazer::stargazer(bayes_table, summary = F)
```

## Supplemental Table S2

```{r}
bayesp1_i = add_ci4int_bayes(p1_inf)[c(9,11,12),]%>% mutate(pattern = "Pattern 1")
bayesp2_i = add_ci4int_bayes(p2_inf)[c(9,11:12),]%>% mutate(pattern = "Pattern 2")
bayesp1_h = add_ci4int_bayes(p1_hybrid)[c(9,11,12),]%>% mutate(pattern = "Pattern 1")
bayesp2_h = add_ci4int_bayes(p2_hybrid)[c(9,11:12),]%>% mutate(pattern = "Pattern 2")

tabbayes_i = bind_rows(bayesp1_i, bayesp2_i)[,c(1,7, 5, 9:10)] %>% as_tibble() %>% 
            rename(term = beta, estimate = `50%`) %>% 
            rename_all(~str_replace(., "%", " %")) %>% 
            mutate(model = "Weakly informative")

tabbayes_h = bind_rows(bayesp1_h, bayesp2_h)[,c(1,7, 5, 9:10)] %>% as_tibble() %>% 
            rename(term = beta, estimate = `50%`) %>% 
            rename_all(~str_replace(., "%", " %")) %>% 
            mutate(model = "Weakly/non- informative")

supp_table = bind_rows(tabbayes_i, tabbayes_h) %>% dplyr::select(model, pattern, everything()) %>% 
  arrange(model, pattern, term) %>% 
  mutate_if(is.numeric, round, 1) %>% 
  mutate(`95% Confidence Interval` = str_c("(", `2.5 %`, ", ", `97.5 %`, ")")) %>% 
  mutate(term = case_when(term == "sex*p" | grepl("SEX", term) ~ " interaction term",
                          TRUE ~ term),
         term = str_remove(term, "(P1|P2|p)"),
         term = str_c(pattern, term)) %>%  
    dplyr::select(-`2.5 %`, -`97.5 %`, -pattern)

supp_table

# stargazer::stargazer(supp_table[,3:4], summary = F)
```
